Hydras \& Co.

### 2022-10-17 doc/chapter-fol.tex
more comments on FOL

Let us give two examples of first-order formulas.

\begin{itemize}
\item $\forall\,v_0\;v_1,\, v_0<v_1\, \longrightarrow\, v_0<v_1 + v_0$
\inputsnippets{fol_Examples/f1Example}
\item $\forall\,v_0,\,\exists\,v_1,\,v_0<v_1$
\inputsnippets{fol_Examples/f2Example}

\end{itemize}

@@

\subsubsection{Depth of a formula}

\inputsnippets{fol/depthDef}

\inputsnippets{fol_Examples/ltdepth1}

@@

@@

>>>>

Let us give a few examples of first-order formulas.

\label{fol:examplesf1f2f3}
\begin{enumerate}
\item $\forall\,v_0, v_0=0 \vee \exists\;v_1,\, v_0= S\;v_1$
\item $\exists\,v_2,\, 0< v_2 \wedge 4=v_2+v_2$
\item $ v_0=0 \vee \exists\;v_1,\, v_0= S\;v_1$
\item $v_0= v_1 + v_1 \leftrightarrow v_0 = v_1 \times 2$
\end{enumerate}
\inputsnippets{fol_Examples/f1Example}
\inputsnippets{fol_Examples/f2Example}

\begin{todo}
Comment on the use of boths list and sets of formulas.
(explained in Russel's paper, p.7).

\inputsnippets{fol/SystemDef}

\end{todo}

@@

The purpose of this section is to present ways of proving 
properties of any formula $a$ (or defining functions over the type of formulas).

Structural induction/recursion principles are generated by 
\texttt{Formula}'s definition, for instance:

\inputsnippets{fol_Examples/FormulaRect}

\subsubsection{Depth of a formula}
As explained in~\cite{OConnor05}, some functions on the \texttt{Formula} type cannot be defined by structural recursion in 
O'Connors formalization. For instance, the substitution of a term to a free variable (because of quantifiers). 
In such cases, we may define functions by well-founded recursion. In this section, we consider the well-founded strict-order associated with a measure: the \emph{depth} of a formula.

\begin{todo}
Forward link to its application.
\end{todo}

The function \texttt{depth} computes the \emph{depth} of the 
 skeleton of any formula.

\inputsnippets{fol/depthDef, fol_Examples/depthCompute}

\subsubsection{Induction on depth}

The function \texttt{depth} allows us to define a strict order on 
\texttt{Formula $L$}.

\inputsnippets{fol/ltDepth, fol_Examples/ltdepth1}

@@

\inputsnippets{fol_examples/depthRecDemo}

@@

The library~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol} contains several induction principles, applied 
throughout \texttt{Ackermann} and \texttt{Goedel} projects.

Let us for instance have a look  at \texttt{Formula\_depth\_ind2}. Its application in order to prove $P\;a$ generates 5 sub-goals.

\inputsnippets{fol_Examples/depthRecDemo2}

\begin{itemize}
\item Goals $1$ to $4$ correspond to  usual proofs by structural induction.
\item Goal $5$ is associated with a universal quantification $f=\forall\,v,a$. In this case, we have to prove that $P\;b$ holds for any formula $b$ which has a depth strictly less than $f$. Such a $b$ may for instance the result of replacing the free occurrences of $v$ in $a$ with any term $t$.
  \begin{todo}
   Make a link to an appropriate example.
  \end{todo}
\end{itemize}

\subsection{Free variables and substitutions}

Section \texttt{Free\_Variables} 
of~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol} defines the list of free variables which occur in a term (respectively in a formula).

We advise the reader to look at these functions, and limit ourselves to give a few examples.

Let us consider again the formulas \texttt{f3} and \texttt{f4} of page~\pageref{fol:examplesf1f2f3}.

The list of free variable occurrences of a formula is given by the function \texttt{freeVarFormula}. The function \texttt{close}
binds all the free variables of a formula with universal quantifiers.

\inputsnippets{fol_Examples/freeVarExamples}
